---
title: "Clustering with easystats"
output: 
  rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{Clustering with easystats}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  eval = TRUE
)

if (!requireNamespace("NbClust", quietly = TRUE) ||
  !requireNamespace("mclust", quietly = TRUE) ||
  !requireNamespace("cluster", quietly = TRUE) ||
  !requireNamespace("ggplot2", quietly = TRUE) ||
  !requireNamespace("see", quietly = TRUE) ||
  !requireNamespace("fpc", quietly = TRUE) ||
  !requireNamespace("performance", quietly = TRUE) ||
  !requireNamespace("dbscan", quietly = TRUE) ||
  !requireNamespace("pvclust", quietly = TRUE)) {
  knitr::opts_chunk$set(eval = FALSE)
} else {
  library(parameters)
  library(NbClust)
  library(performance)
  library(cluster)
  library(fpc)
  library(mclust)
  library(insight)
  library(dbscan)
  library(pvclust)
}
```

This vignette can be referred to by citing the package:

- Lüdecke, D., Ben-Shachar, M. S., Patil, I., \& Makowski, D. (2020). *Extracting, computing and exploring the parameters of statistical models using R*. Journal of Open Source Software, 5(53), 2445. https://doi.org/10.21105/joss.02445



Note that in order to fully use all the methods demonstrated below, you will need to additionally install the packages below:

```{r eval = FALSE}
install.packages(c("NbClust", "mclust", "pvclust", "cluster", "fpc", "dbscan"))
```

## Introduction

Clustering traditionally refers to the identification of groups of observations (i.e., data rows). It differs from methods like [**PCA or Factor Analysis**](https://easystats.github.io/parameters/articles/efa_cfa.html), which are usually applied on variables (i.e., columns). That said, it is possible to *transpose* your data (columns become rows) to apply clustering on variables.

There are many clustering algorithms (see [this for an overview](https://scikit-learn.org/stable/modules/clustering.html)), but they can grouped in two categories: **supervised** and **unsupervised** techniques. In **supervised** techniques, you have to explicitly specify [**how many clusters**](https://easystats.github.io/parameters/reference/n_clusters.html) you want to extract. **Unsupervised** techniques, on the other hand, will estimate this number as part of their algorithm. Note that there are no inherently superior and inferior clustering methods, each come with their sets of limitations and benefits.

As an example in the tutorial below, we will use the **iris** dataset, for which we know that there are 3 "real" clusters (the 3 Species of flowers). Let's first start with visualizing the 3 "real" clusters on a 2D space of the variables created through PCA. 


```{r}
library(ggplot2)
library(parameters)
library(see)

set.seed(33)  # Set random seed

# Select the first 4 numeric columns (drop the Species fator)
data <- iris[1:4]  
head(data)  # Print the 6 first rows

# Run PCA
pca <- principal_components(data, n = 2)
pca_scores <- predict(pca, names = c("PCA_1", "PCA_2"))
pca_scores$True_Clusters <- iris$Species  # Add real clusters

# Visualize
ggplot(pca_scores, aes(x = PCA_1, y = PCA_2, color = True_Clusters)) + 
  geom_point() +
  theme_modern()
```

While the **setosa** species stands out quite clearly in this PCA space, the separation between the two other species appear less clear cut. Let's see how data-driven clustering performs, and if we manage to retrieve these 3 clusters.


## Supervised Clustering Methods

### How Many Clusters to Extract?

There is no easy answer to that important question. The best way is to have strong expectations or hypotheses. If you don't, well, researchers have came up with data-driven solutions to estimate the optimal number of clusters. The problem is that there are now a lot of these numerical methods, and that they don't always agree...

Because there is no clearly better method, we have implemented in *easystats* a consensus-based algorithm that runs many of these methods, and returns the number of clusters that is the most agreed upon.


```{r}
n <- n_clusters(data, package = c("easystats", "NbClust", "mclust"))
n
plot(n)
```

As we can see, most methods suggest the existence of **2 clusters**, followed by a **3-clusters** solution. It seems like the data does not clearly discriminate between the 3 species of flowers. This discrepancy between what is, and what we can recover from real-world data, is a fundamental issue in data science.


### K-Means

We won't go too much into details about the mathematics and intuition behind these clustering methods, as good [resources](https://scikit-learn.org/stable/modules/clustering.html) are available all over the internet. Instead, we'll focus on how to apply them.

K-means is one of the most basic clustering algorithm, available in base R through the `kmeans()` function. However, we provide in easystats a unified function to run different clustering algorithms: [**cluster_analysis()**](https://easystats.github.io/parameters/reference/cluster_analysis.html). *(Note that k-means is a non-deterministic algorithm; running it multiple times will result in different results!)*

Now that we know how many clusters we want to extract (let's say that we have a strong hypothesis on 3, which is partially supported by the consensus method for estimating the optimal number of clusters).


```{r}
rez_kmeans <- cluster_analysis(data, n = 3, method = "kmeans")

rez_kmeans  # Show results
```

One can extract the cluster assignments to use it as a new variable by using `predict()`.

```{r}
predict(rez_kmeans)  # Get clusters
```

It is also possible to visualize the clustering on a 2D space created by the two first PCA components:

```{r}
plot(rez_kmeans) +
  theme_modern()
```

### Hierarchical Clustering

Hierarchical clustering is also a common clustering algorithm, available in base R through the `hclust()` function. This method is a bit different in the sense that is does not straight up return clusters. Instead, in creates a hierarchical structure (a *dendrogram*), a tree from which we can *cut* branches to get a given number of clusters. Note that this "tree" cutting can be done in an unsupervised fashion too using bootstrapping (which we will apply in the next section).

```{r}
rez_hclust <- cluster_analysis(data, n = 3, method = "hclust")

rez_hclust  # Show results

# Visualize
plot(rez_hclust) + theme_modern()  # Visualize 
```

### Hierarchical K-Means

Hierarchical K-Means, as its name suggest, is essentially a combination of K-Means and hierarchical clustering that aims at improving the stability and robustness of the results.

```{r}
rez_hkmeans <- cluster_analysis(data, n = 3, method = "hkmeans")

rez_hkmeans  # Show results

# Visualize
plot(rez_hkmeans) + theme_modern()  # Visualize 
```

### K-Medoids (PAM)

Clustering around "medoids", instead of "centroid", is considered to be a more robust version of K-means. See `cluster::pam()` for more information. 

```{r}
rez_pam <- cluster_analysis(data, n = 3, method = "pam")

rez_pam  # Show results

# Visualize
plot(rez_pam) + theme_modern()  # Visualize 
```


## Unsupervised Clustering Methods

Unsupervised clustering methods estimate the optimal number of clusters themselves (hence, `n = NULL` as we don't pre-specify a given number of clusters). Note that unsupervised methods can sometimes identify observations that do not fit under any clusters (i.e., **"outliers"**). They will be classified as belonging to the cluster "0" (which is not a real cluster, but rather groups all the outliers).

### Bootstrapped Hierarchical Clustering

This method computes p-values for each cluster of the hierarchical cluster structure, and returns the **significant** clusters. This method can return a larger number of smaller clusters and, because it's based on bootstrapping, is quite slow.

```{r}
rez_hclust2 <- cluster_analysis(data, 
                        n = NULL, 
                        method = "hclust", 
                        iterations = 500,
                        ci = 0.90)

rez_hclust2  # Show results
plot(rez_hclust2) + theme_modern()  # Visualize
```


### DBSCAN

Although the DBSCAN method is quite powerful to identify clusters, it is highly dependent on its parameters, namely, `eps` and the `min_size`. Regarding the latter, the minimum size of any cluster is set by default to `0.1` (i.e., 10\% of rows), which is appropriate to avoid having too small clusters. 

The "optimal" **eps** value can be estimated using the [`n_clusters_dbscan()`](https://easystats.github.io/parameters/reference/cluster_analysis.html) function:

```{r}
eps <- n_clusters_dbscan(data, min_size = 0.1) 
eps
plot(eps)
```

It seems like the numeric method to find the elbow of the curve doesn't work well, and returns a value that is too high. Based on visual assessment, the elbow seems to be located around `eps = 1.45`.

```{r}
rez_dbscan <- cluster_analysis(data, method = "dbscan", dbscan_eps = 1.45)

rez_dbscan  # Show results
plot(rez_dbscan) + theme_modern()  # Visualize
```

### Hierarchical K-Means

Hierarchical DBSCAN is a variant that does not require the critical **EPS** argument. It computes the hierarchy of all DBSCAN solutions, and then finds the optimal cuts in the hierarchy using a stability-based extraction method.

```{r}
rez_hdbscan <- cluster_analysis(data, method = "hdbscan")

rez_hdbscan  # Show results

# Visualize
plot(rez_hdbscan) + theme_modern()  # Visualize 
```


### K-Medoids with estimation of number of clusters (pamk)

This is K-Medoids with an integrated estimation of the number of clusters. See `fpc::pamk` for more details.

```{r}
rez_pamk <- cluster_analysis(data, method = "pamk")

rez_pamk  # Show results

# Visualize
plot(rez_pamk) + theme_modern()  # Visualize 
```

### Mixture

Model-based clustering based on finite Gaussian mixture models. Models are estimated by EM algorithm initialized by hierarchical model-based agglomerative clustering. The optimal model is then selected according to BIC.

```{r}
library(mclust)

rez_mixture <- cluster_analysis(data, method = "mixture")

rez_mixture  # Show results

# Visualize
plot(rez_mixture) + theme_modern()  # Visualize 
```

## Metaclustering

One of the core "issue" of statistical clustering is that, in many cases, different methods will give different results. The **metaclustering** approach proposed by *easystats* (that finds echoes in *consensus clustering*; see Monti et al., 2003) consists of treating the unique clustering solutions as a ensemble, from which we can derive a probability matrix. This matrix contains, for each pair of observations, the probability of being in the same cluster. For instance, if the 6th and the 9th row of a dataframe has been assigned to a similar cluster by 5 our of 10 clustering methods, then its probability of being grouped together is 0.5.

Metaclustering is based on the hypothesis that, as each clustering algorithm embodies a different prism by which it sees the data, running an infinite amount of algorithms would result in the emergence of the "true" clusters. As the number of algorithms and parameters is finite, the probabilistic perspective is a useful proxy. This method is interesting where there is no obvious reasons to prefer one over another clustering method, as well as to investigate how robust some clusters are under different algorithms.


```{r}
list_of_results <- list(rez_kmeans, rez_hclust, rez_hkmeans, rez_pam,
                        rez_hclust2, rez_dbscan, rez_hdbscan, rez_mixture)

probability_matrix <- cluster_meta(list_of_results)

# Plot the matrix as a reordered heatmap
heatmap(probability_matrix, scale = "none", 
        col = grDevices::hcl.colors(256, palette = "inferno"))
```



The dendrogram (which is a **hierarchical clustering of the clustering solution**, hence the name of **meta**clustering), as well as the heatmap (in which the darker squares represent a higher probability of belonging to the same cluster) shows that there is one metacluster consisting of the 1-50 first rows (bottom left), and then the rest of the observations are closer to one another. However, two subclusters are still visible, corresponding to the "true" species.

The metaclustering approach confirms our initial hypothesis, *the **setosa** species stands out quite clearly, and the separation between the two other species is less clear cut*.

## Resources

- [Clustering algorithms overview](https://scikit-learn.org/stable/modules/clustering.html)
- [Density-based Clustering](https://www.datanovia.com/en/lessons/dbscan-density-based-clustering-essentials/)