---
title: "Analysing Longitudinal or Panel Data"
output: 
  github_document:
    toc: true
    fig_width: 10.08
    fig_height: 6
  rmarkdown::html_vignette:
    toc: true
    fig_width: 10.08
    fig_height: 6
vignette: >
  %\VignetteIndexEntry{Analysing Longitudinal or Panel Data}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE, 
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  eval = TRUE
)

if (!requireNamespace("lme4", quietly = TRUE) ||
    !requireNamespace("lfe", quietly = TRUE)) {
  knitr::opts_chunk$set(eval = FALSE)
} else {
  library(parameters)
  library(lme4)
  library(lfe)
}

set.seed(333)
```

This vignette explains the rational behind the [`demean()`](https://easystats.github.io/parameters/reference/demean.html) function. We give recommendations how to analyze multilevel or hierarchical data structures, when macro-indicators (or level-2 predictors, or higher-level units, or more general: _group-level predictors_) are used as covariates and the model suffers from **heterogeneity bias** [@bell_explaining_2015].

# Sampel data used in this vignette

```{r}
library(parameters)
data("qol_cancer")
```

* Variables:
  * `QoL`      : Response (quality of life of patient)
  * `phq4`     : Patient Health Questionnaire, **time-varying** variable
  * `hospital` : Location of treatment, **time-invariant** variable, co-variate
  * `education`: Educational level, **time-invariant** variable, co-variate
  * `ID`       : patient ID
  * `time`     : time-point of measurement

# Heterogeneity bias

Heterogeneity bias occurs when group-level predictors vary within and across groups, and hence
fixed effects may correlate with group (or random) effects. This is a typical situation when analyzing longitudinal or panel data: Due to the repeated measurements of persons, the "person" (or subject-ID) is now a level-2 variable. Predictors at level-1 ("fixed effects"), e.g.  self-rated health or income, now have an effect at level-1 ("within"-effect) and at higher-level units (level-2, the subject-level, which is the "between"-effect). This inevitably leads to correlating fixed effects and error terms. 

# Adressing heterogeneity bias: the Fixed Effects Regression (FE) approach

Fixed effects regression models (FE) are a popular approach for panel data analysis in particular in econometrics and considered as gold standard. To avoid the problem of heterogeneity bias, in FE all higher-level variance (and thus, any between-effects), "are controlled out using the higher-level entities themselves, included in the model as dummy variables" [@bell_explaining_2015]. As a consequence, FE models are only able estimate _within-effects_.

To remove between-effects and only model within-effects, the data needs some preparation: _de-meaning_. De-meaning, or _person-mean centering_, or _centering within clusters_, takes away the higher-level mean from the regression equation, and as such, FE avoids estimating a parameter for each higher-level unit.

## Computing the de-meaned and group-meaned variables

```{r}
qol_cancer <- cbind(
  qol_cancer,
  demean(qol_cancer, select = c("phq4", "QoL"), group = "ID")
)
```

Now we have:

  * `phq4_between`: time-varying variable with the mean of `phq4` across all time-points, for each patient (ID).
  * `phq4_within`: the de-meaned time-varying variable `phq4`.

A FE model is a classical linear model, where

 * Intercept is removed
  * time-invariant predictors are not allowed to be included
  * the group-level factor is included as predictor
  * time-varying predictors are de-meaned ("person-mean centered", indicating the "within-subject" effect)

```{r}
fe_model1 <- lm(
  QoL ~ 0 + time + phq4_within + ID,
   data = qol_cancer
)
# we use only the first two rows, because the remaining rows are
# the estimates for "ID", which is not of interest here...
model_parameters(fe_model1)[1:2, ]


# instead of removing the intercept, we could also use the 
# de-meaned response...
fe_model2 <- lm(
  QoL_within ~ time + phq4_within + ID,
   data = qol_cancer
)
model_parameters(fe_model2)[2:3, ]


# we compare the results with those from the "lfe"-package for panel data
library(lfe)
fe_model3 <- felm(
  QoL ~ time + phq4_within | ID,
   data = qol_cancer
)
model_parameters(fe_model3)
```

As we can see, the _within-effect_ of PHQ-4 is `-3.66`, hence the mean of the change for an average individual case in our sample (or, the "net" effect), is `-3.66`.

But what about the between-effect? How do people with higher PHQ-4 score differ from people with lower PHQ-4 score? Or what about educational inequalities? Do higher educated people have a higher PHQ-4 score than lower educated people?

This question cannot be answered with FE regression. But: "Can one fit a multilevel model with varying intercepts (or coefficients) when the units and predictors correlate? The answer is yes. And the solution is simple." [@bafumi_fitting_2006]

# Adressing heterogeneity bias: the Mixed Model approach

Mixed models include different levels of sources of variability (i.e. error terms at each level). Predictors used at level-1 that are varying across higher-level units will thus have residual errors at both level-1 and higher-level units. "Such covariates contain two parts: one that is specific to the higher-level entity that does not vary between occasions, and one that represents the difference between occasions, within higher-level entities" [@bell_explaining_2015]. Hence, the error terms will be correlated with the covariate, which violates one of the assumptions of mixed models (iid, independent and identically distributed error terms) - also known and described above as _heterogeneity bias_.

But how can this issue be addressed outside the FE framework?

There are several ways how to address this using a mixed models approach: 

  * Correlated group factors and predictors are no problem anyway, because [partial pooling](https://www.tjmahr.com/plotting-partial-pooling-in-mixed-effects-models/) allows estimates of units o borrow strength from the whole sample and shrink toward a common mean (@shor_bayesian_2007).
  * If predictor and group factors correlate, one can remove this correlation by group-meaning [or "mean within clusters", @bafumi_fitting_2006; @gelman_data_2007, Chap. 12.6.].
  * When time-varying predictors are "decomposed" into their time-varying and time-invariant components (demeaning), then mixed models can model **both** within- and between-subject effects [@bell_fixed_2018] - this approach is essentially a further development of a long-known recommendation by Mundlak [@mundlak_pooling_1978].

For now, we will follow the last recommendation and use the within- and between-version of `phq4`.

```{r}
library(lme4)
mixed_1 <- lmer(
  QoL ~ time + phq4_within + phq4_between + (1 | ID),
   data = qol_cancer
)
model_parameters(mixed_1)

# compare to FE-model
model_parameters(fe_model1)[1:2, ]
```

As we can see, the estimates and standard errors are identical. The argument _against_ the use of mixed models, i.e. that using mixed models for panel data will yield biased estimates and standard errors, is based on an incorrect model specification [@mundlak_pooling_1978]. As such, when the (mixed) model is properly specified, the estimator of the mixed model is identical to the 'within' (i.e. FE) estimator.

As a consequence, we cannot only use the above specified mixed model for panel data, we can even specify more complex models including within-effects, between-effects or random effects variation. A mixed models approach can model the causes of endogeneity explicitly by including separating the within- and between-effects of time-varying fixed effects and including time-constant fixed effects.

```{r}
mixed_2 <- lmer(
  QoL ~ time + phq4_within + phq4_between + education + (1 + time | ID),
   data = qol_cancer
)
model_parameters(mixed_2)
```

For more complex models, within-effects will naturally change slightly and are no longer identical to simpler FE models. This is no "bias", but rather the result of building more complex models: FE models lack information of variation in the group-effects or between-subject effects. Furthermore, FE models cannot include random slopes, which means that fixed effects regressions are neglecting "cross-cluster differences in the effects of lower-level controls (which) reduces the precision of estimated context effects, resulting in (...) low statistical power" [@heisig_costs_2017].

# Conclusion: Complex Random Effects Within-Between Models

Depending on the structure of the data, the best approach to analyzing panel data would be a so called "complex random effects within-between" model [@bell_fixed_2018]:

```{r echo=FALSE}
f <- "y<sub>it</sub> = &beta;<sub>0</sub> + &beta;<sub>1W</sub> (x<sub>it</sub> - &#x035E;x<sub>i</sub>) + &beta;<sub>2B</sub> &#x035E;x<sub>i</sub> + &beta;<sub>3</sub> z<sub>i</sub> + &upsilon;<sub>i0</sub> + &upsilon;<sub>i1</sub> (x<sub>it</sub> - &#x035E;x<sub>i</sub>) + &epsilon;<sub>it</sub>"
knitr::asis_output(f)
```

```{r echo=FALSE}
f <- "<ul><li>x<sub>it</sub> - &#x035E;x<sub>i</sub> is the de-meaned predictor, <em>phq4_within</em></li><li>&#x035E;x<sub>i</sub> is the group-meaned predictor, <em>phq4_between</em></li><li>&beta;<sub>1W</sub> is the coefficient for phq4_within (within-subject)</li><li>&beta;<sub>2B</sub> is the coefficient for phq4_between (bewteen-subject)</li><li>&beta;<sub>3</sub> is the coefficient for time-constant predictors, such as `hospital` or `education` (bewteen-subject)</li></ul>"
knitr::asis_output(f)
```

Written in R-code, the model is written down like this:

```{r}
rewb <- lmer(
  QoL ~ time + phq4_within + phq4_between + education + 
    (1 + time | ID) + (1 + phq4_within | ID),
   data = qol_cancer
)
```

The benefit of this kind of model is that you have information on within-, between- and other time-constant (i.e. between) effects...

```{r}
model_parameters(rewb)
```

... but you can also model the variation of (group) effects across time (and probably space), and you can even include more higher-level units (e.g. nested design or cross-classified design with more than two levels):

```{r}
random_parameters(rewb)
```

# References

