---
title: "Formatting Model Parameters"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_width: 10.08
    fig_height: 6
tags: [r, parameters, table layout]
vignette: >
  %\VignetteIndexEntry{Printing Model Parameters}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
---

```{r , include=FALSE}
library(knitr)
options(knitr.kable.NA = "")
options(digits = 2)

knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  out.width = "100%",
  tidy.opts = list(width.cutoff = 100)
)

if (!requireNamespace("gt", quietly = TRUE) ||
  !requireNamespace("magrittr", quietly = TRUE) ||
  !requireNamespace("glmmTMB", quietly = TRUE)) {
  knitr::opts_chunk$set(eval = FALSE)
} else {
  library(parameters)
  library(gt)
  library(magrittr)
  library(insight)
}

set.seed(333)
```

`model_parameters()` and `compare_parameters()` are functions that return a data frame of model summaries in a consistent way. The printed table of those summaries is formatted to make the output more readable and removes or collapses redundant columns, to get a compact and yet comprehensive summary table. _(N.B. for developers: the function [`standardize_names()`](https://easystats.github.io/insight/reference/standardize_names.html) standardizes the column names, so column names are consistent and the same for any model object, also in **broom** style, which makes it easy to build your packages on top of the **parameters** package.)_

The default [print-methods](https://easystats.github.io/parameters/reference/print.parameters_model.html) for `model_parameters()` and `compare_parameters()` allows the user to modify the layout and style of the output.

# Summaries for a single model

In the following examples for `model_parameters()`, which returns tabular output for single models, are shown.

## Pretty parameter names formatting

By default, the argument `pretty_names` is `TRUE`, meaning that parameter names are formatted to make them more "human readable", i.e. factor levels are separated from the variable names, interactions are denoted by `*` etc.

```{r}
library(parameters)
data(iris)
model <- lm(Sepal.Length ~ Species * Petal.Length, data = iris)
model_parameters(model)

mp <- model_parameters(model)
print(mp, pretty_names = FALSE)
```

## Splitting model components

Again by default, the argument `split_components` is `TRUE`, which means that models with multiple components like fixed and random effects, count and zero-inflated part etc. are split into separate tables in the output.

```{r}
library(glmmTMB)
data("Salamanders")
model <- glmmTMB(count ~ spp + mined + (1 | site),
                 ziformula = ~spp + mined,
                 family = nbinom2(), 
                 data = Salamanders)
model_parameters(model)
```

Redundant columns are removed. The related model component is shown as table header. However, you can also return a single table:

```{r}
mp <- model_parameters(model)
print(mp, split_component = FALSE)
```

## Adding model summaries

A model summary can be added to the table when `summary = TRUE` in the call to `model_parameters()`:

```{r}
model <- lm(Sepal.Length ~ Species * Petal.Length, data = iris)
model_parameters(model, summary = TRUE)
```

## Changing number of digits

`digits` changes the digits for coefficients, standard errors and statistics. `ci_digits` and `p_digits` are especially for the confidence intervals and p-values.

```{r}
model <- lm(Sepal.Length ~ Species, data = iris)
model_parameters(model, digits = 4)
```

p-values can be displayed in exact, scientific notation if required.

```{r}
model_parameters(model, p_digits = "scientific")
```

## Group parameters

The `groups` argument can be used to group parameters in the table. `groups` must be a named list, where the names of the list elements equal the header of each group, while the values of the list elements equal the parameter names, or the position of the parameters in the table (data frame).

In the following example, we see the names of the parameters in the `Parameter` column, while the rownumbers indicate their position.

```{r}
data(mtcars)
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$gear <- as.factor(mtcars$gear)
model <- lm(mpg ~ hp + gear * vs + cyl + drat, data = mtcars)

# don't select "Intercept" parameter
mp <- model_parameters(model, parameters = "^(?!\\(Intercept)")

# inspect data frame
as.data.frame(mp)
```

Now we create a group named `"Engine"`, which encompasses the parameters `"cyl6"`, `"cyl8"`, `"vs"` and `"hp"`. The `"Interactions"` group includes `"gear4:vs"` and `"gear5:vs"`. The group `"controls"` has the parameters from rows 2, 3 and 7.

Note that the parameters in the table summary are re-ordered according to the order specified in `groups`.

```{r}
# group parameters, either by parameter name or position
print(mp, groups = list("Engine" = c("cyl6", "cyl8", "vs", "hp"),
                        "Interactions" = c("gear4:vs", "gear5:vs"), 
                        "Controls" = c(2, 3, 7))) # gear 4 and 5, drat
```

# Summaries for multiple models

`compare_parameters()` (or its alias `compare_models()`) allows to create tables for multiple models, aligned side by side.

By default, estimates and confidence intervals are shown.

```{r}
data(iris)
lm1 <- lm(Sepal.Length ~ Species, data = iris)
lm2 <- lm(Sepal.Length ~ Species + Petal.Length, data = iris)
lm3 <- lm(Sepal.Length ~ Species * Petal.Length, data = iris)
compare_parameters(lm1, lm2, lm3)
```

## Changing style of column output

By default, estimates and confidence intervals are shown. Using `style` allows us to create different output, e.g. standard errors instead of confidence intervals, or including p-values.

```{r}
compare_parameters(lm1, lm2, lm3, style = "se_p")
```

## Defining column names

The column names for the models are by default the objects' names. You can define own names using the `column_names` argument.

```{r}
compare_parameters(
  lm1, lm2, lm3,
  column_names = c("First Model", "Second Model", "Third Model")
)
```

## Group parameters of multiple model tables

Grouping parameters works for `compare_models()` in the same way as shown above for `model_parameters()`.

```{r}
lm1 <- lm(Sepal.Length ~ Species + Petal.Length, data = iris)
lm2 <- lm(Sepal.Width ~ Species * Petal.Length, data = iris)

# remove intercept
cp <- compare_parameters(lm1, lm2, parameters = "^(?!\\(Intercept)")

# look at parameters names, to know their names for "groups" argument
as.data.frame(cp)$Parameter

# create groups. Interactions only present in 2nd model
print(cp, groups = list(Species = c("Species (versicolor)", 
                                    "Species (virginica)"),
                        Interactions = c("Species (versicolor) * Petal.Length",
                                         "Species (virginica) * Petal.Length"),
                        Controls = "Petal.Length"))
```

## More advances tables and markdown / HTML formatting

The `print_md()` as well as `print_html()` functions can be used to create markdown (for knitting to PDF or Word) and HTML tables. 

Meanwhile, there are a lot of additional packages that allow users to have even more flexibility regarding table layouts. One package we can recommend is the [*modelsummary* package](https://vincentarelbundock.github.io/modelsummary/).
