---
title: "Summary of Model Parameters"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_width: 10.08
    fig_height: 6
tags: [r, parameters, variable selection, feature selection]
vignette: >
  %\VignetteIndexEntry{Summary of Model Parameters}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
options(knitr.kable.NA = "")
options(digits = 2)

knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  out.width = "100%",
  tidy.opts = list(width.cutoff = 120)
)

if (!requireNamespace("poorman", quietly = TRUE) ||
  !requireNamespace("effectsize", quietly = TRUE) ||
  !requireNamespace("BayesFactor", quietly = TRUE) ||
  !requireNamespace("lme4", quietly = TRUE) ||
  !requireNamespace("metafor", quietly = TRUE) ||
  !requireNamespace("lavaan", quietly = TRUE) ||
  !requireNamespace("nFactors", quietly = TRUE) ||
  !requireNamespace("EGAnet", quietly = TRUE) ||
  !requireNamespace("brms", quietly = TRUE) ||
  !requireNamespace("psych", quietly = TRUE) ||
  !requireNamespace("rstanarm", quietly = TRUE) ||
  !requireNamespace("glmmTMB", quietly = TRUE) ||
  !requireNamespace("GLMMadaptive", quietly = TRUE) ||
  !requireNamespace("FactoMineR", quietly = TRUE)) {
  knitr::opts_chunk$set(eval = FALSE)
} else {
  library(parameters)
  library(poorman)
  library(effectsize)
  library(EGAnet)
  library(psych)
  library(nFactors)
  library(brms)
  library(GLMMadaptive)
  library(FactoMineR)
  
}

set.seed(333)
```

The `model_parameters()` function (also accessible via the shortcut
`parameters()`) allows you to extract the parameters and their characteristics
from various models in a consistent way. It can be considered as a lightweight
alternative to [`broom::tidy()`](https://github.com/tidymodels/broom), with some
notable differences:

- The names of the returned data frame are **specific** to their content. For
  instance, the column containing the statistic is named following the statistic
  name, i.e., *t*, *z*, etc., instead of a generic name such as *statistic*
  (**however**, you can get standardized (generic) column names using
  [`standardize_names()`](https://easystats.github.io/insight/reference/standardize_names.html)).

- It is able to compute or extract indices not available by default, such as
  **p*-values**, **CIs**, etc.

- It includes **feature engineering** capabilities, including parameters
  [**bootstrapping**](https://easystats.github.io/parameters/reference/bootstrap_model.html).

## Correlations and *t*-tests

### Frequentist

```{r}
cor.test(iris$Sepal.Length, iris$Sepal.Width) %>%
  parameters()
```

```{r}
t.test(mpg ~ vs, data = mtcars) %>%
  parameters()
```


### Bayesian

```{r}
library(BayesFactor)

BayesFactor::correlationBF(iris$Sepal.Length, iris$Sepal.Width) %>%
  parameters()
```

```{r}
BayesFactor::ttestBF(formula = mpg ~ vs, data = mtcars) %>%
  parameters()
```

## ANOVAs

Indices of effect size for ANOVAs, such as partial and non-partial versions of
`eta_squared()`, `epsilon_sqared()` or `omega_squared()` are powered by the
[**effectsize**-package](https://easystats.github.io/effectsize/). However,
**parameters** uses these function to compute such indices for parameters
summaries, including confidence intervals

### Simple

```{r}
aov(Sepal.Length ~ Species, data = iris) %>%
  parameters(
    omega_squared = "partial",
    eta_squared = "partial",
    epsilon_squared = "partial"
  )
```

Let's complicate things further with an interaction term:

```{r}
aov(Sepal.Length ~ Species * Sepal.Width, data = iris) %>%
  parameters(
    omega_squared = "partial",
    eta_squared = "partial",
    ci = .8
  )
```

### Repeated measures

`parameters()` (resp. its alias `model_parameters()`) also works on repeated
measures ANOVAs, whether computed from `aov()` or from a mixed model.

```{r}
aov(mpg ~ am + Error(gear), data = mtcars) %>%
  parameters()
```

## Regressions (GLMs, Mixed Models, GAMs, ...)

`parameters()` (resp. its alias `model_parameters()`) was mainly built with
regression models in mind. It works for many types of models and packages,
including mixed models and Bayesian models.

### GLMs

```{r}
glm(vs ~ poly(mpg, 2) + cyl, data = mtcars, family = binomial()) %>%
  parameters()
```

```{r}
# show Odds Ratios and Wald-method for degrees of freedom
glm(vs ~ poly(mpg, 2) + cyl, data = mtcars, family = binomial()) %>%
  parameters(exponentiate = TRUE, df_method = "wald")
```

```{r}
# show Odds Ratios and include model summary
glm(vs ~ poly(mpg, 2) + cyl, data = mtcars, family = binomial()) %>%
  parameters(exponentiate = TRUE, summary = TRUE)
```

### Mixed Models

```{r}
library(lme4)

lmer(Sepal.Width ~ Petal.Length + (1 | Species), data = iris) %>%
  parameters()
```

### Mixed Models, without Random Effects Variances

```{r}
lmer(Sepal.Width ~ Petal.Length + (1 | Species), data = iris) %>%
  parameters(effects = "fixed")
```

### Mixed Model with Zero-Inflation Model

```{r}
library(GLMMadaptive)
library(glmmTMB)
data("Salamanders")
model <- mixed_model(
  count ~ spp + mined,
  random = ~ 1 | site,
  zi_fixed = ~ spp + mined,
  family = zi.negative.binomial(),
  data = Salamanders
)
parameters(model)
```

### Mixed Models with Dispersion Model

```{r}
library(glmmTMB)

sim1 <- function(nfac = 40, nt = 100, facsd = 0.1, tsd = 0.15, mu = 0, residsd = 1) {
  dat <- expand.grid(fac = factor(letters[1:nfac]), t = 1:nt)
  n <- nrow(dat)
  dat$REfac <- rnorm(nfac, sd = facsd)[dat$fac]
  dat$REt <- rnorm(nt, sd = tsd)[dat$t]
  dat$x <- rnorm(n, mean = mu, sd = residsd) + dat$REfac + dat$REt
  dat
}

set.seed(101)
d1 <- sim1(mu = 100, residsd = 10)
d2 <- sim1(mu = 200, residsd = 5)
d1$sd <- "ten"
d2$sd <- "five"
dat <- rbind(d1, d2)
model <- glmmTMB(x ~ sd + (1 | t), dispformula = ~sd, data = dat)

parameters(model)
```

### Bayesian Models

`model_parameters()` also works with Bayesian models from the **rstanarm**
package:

```{r}
library(rstanarm)

# if you are unfamiliar with the `refresh` argument here, it just avoids
# printing few messages to the console
stan_glm(mpg ~ wt * cyl, data = mtcars, refresh = 0) %>%
  parameters()
```

Additionally, it also works for models from the **brms** package.

For more complex models, specific model components can be printed using the
arguments `effects` and `component` arguments.

```{r}
library(brms)
data(fish)
set.seed(123)

# fitting a model using `brms`
model <- brm(
  bf(
    count ~ persons + child + camper + (1 | persons),
    zi ~ child + camper + (1 | persons)
  ),
  data = fish,
  family = zero_inflated_poisson(),
  refresh = 0
)

parameters(model, component = "conditional")

parameters(model, effects = "all", component = "all")
```

To include information about the random effect parameters (group levels), set
`group_level = TRUE`:

```{r}
parameters(model, effects = "all", component = "conditional", group_level = TRUE)
```

## Structural Models (PCA, EFA, CFA, SEM...)

The **parameters** package extends the support to structural models.

### Principal Component Analysis (PCA) and Exploratory Factor Analysis (EFA) 

```{r}
library(psych)

psych::pca(mtcars, nfactors = 3) %>%
  parameters()
```

We will avoid displaying a graph while carrying out factor analysis:

```{r}
library(FactoMineR)

FactoMineR::FAMD(iris, ncp = 3, graph = FALSE) %>%
  parameters()
```

### Confirmatory Factor Analysis (CFA) and Structural Equation Models (SEM)

#### Frequentist

```{r}
library(lavaan)

model <- lavaan::cfa(" visual  =~ x1 + x2 + x3
                       textual =~ x4 + x5 + x6
                       speed   =~ x7 + x8 + x9 ",
  data = HolzingerSwineford1939
)

model_parameters(model)
```

#### Bayesian

`blavaan` to be done.

## Meta-Analysis

`parameters()` also works for `rma`-objects from the **metafor** package.

```{r}
library(metafor)

mydat <- data.frame(
  effectsize = c(-0.393, 0.675, 0.282, -1.398),
  standarderror = c(0.317, 0.317, 0.13, 0.36)
)

rma(yi = effectsize, sei = standarderror, method = "REML", data = mydat) %>%
  model_parameters()
```

## Plotting Model Parameters

There is a `plot()`-method implemented in the
[**see**-package](https://easystats.github.io/see/). Several examples are shown
[in this vignette](https://easystats.github.io/see/articles/parameters.html).
