---
title: "Printing Model Parameters"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_width: 10.08
    fig_height: 6
tags: [r, parameters, table layout]
vignette: >
  %\VignetteIndexEntry{Printing Model Parameters}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
---

```{r, include=FALSE}
options(knitr.kable.NA = "")
options(digits = 2)

knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = requireNamespace("gt", quietly = TRUE),
  comment = "#>",
  out.width = "100%",
  tidy.opts = list(width.cutoff = 100)
)

pkgs <- c("gt", "glmmTMB")
successfully_loaded <- sapply(pkgs, requireNamespace, quietly = TRUE)

if (all(successfully_loaded)) {
  library(parameters)
  library(datawizard)
  library(glmmTMB)
  library(gt)
}

set.seed(333)
```

`model_parameters()` and `compare_parameters()` are functions that return a data frame of model summaries in a consistent way. The printed table of those summaries is formatted to make the output more readable and removes or collapses redundant columns, to get a compact and yet comprehensive summary table. _(N.B. for developers: the function [`standardize_names()`](https://easystats.github.io/insight/reference/standardize_names.html) standardizes the column names, so column names are consistent and the same for any model object, also in **broom** style, which makes it easy to build your packages on top of the **parameters** package.)_

The default [print-methods](https://easystats.github.io/parameters/reference/print.parameters_model.html) for `model_parameters()` and `compare_parameters()` allows the user to modify the layout and style of the output.

# Summaries for a single model

In the following examples for `model_parameters()`, which returns tabular output for single models, are shown.

## Pretty parameter names formatting

By default, the argument `pretty_names` is `TRUE`, meaning that parameter names are formatted to make them more "human readable", i.e. factor levels are separated from the variable names, interactions are denoted by `*` etc.

```{r}
library(parameters)
data(iris)
model <- lm(Sepal.Length ~ Species * Petal.Length, data = iris)
model_parameters(model)

mp <- model_parameters(model)
print(mp, pretty_names = FALSE)
```

If data is [labelled](https://strengejacke.github.io/sjlabelled/articles/intro_sjlabelled.html), `pretty_names = "labels"` will use variable and value labels as pretty names. If data is not labelled, default pretty names will be used.

```{r eval=successfully_loaded["datawizard"]}
data(efc, package = "datawizard")
model <- lm(neg_c_7 ~ e42dep + c172code, data = efc)

# default printing
model_parameters(model)

# using value and variable labels
mp <- model_parameters(model)
print(mp, pretty_names = "labels")
```

## More compact output

Using `summary`, or the `select` argument via the `print()` method allows for a more compact table, in case not all information is required. `summary()` will return the coefficient, confidence intervals and p-values. `select` allows to select specific columns only.

```{r}
data(iris)
model <- lm(Sepal.Length ~ Species * Petal.Length, data = iris)

result <- model_parameters(model)

# Coefficients, CI and p
summary(result)

# Parameter name, SE and p
print(result, select = c("Parameter", "SE", "p"))
```

The `select` argument can also be used for a more customized output. See examples below for `compare_parameters()`.

## Splitting model components

Again by default, the argument `split_components` is `TRUE`, which means that models with multiple components like fixed and random effects, count and zero-inflated part etc. are split into separate tables in the output.

```{r eval=successfully_loaded["glmmTMB"]}
library(glmmTMB)
data("Salamanders")
model <- glmmTMB(count ~ spp + mined + (1 | site),
  ziformula = ~ spp + mined,
  family = nbinom2(),
  data = Salamanders
)
model_parameters(model)
```

Redundant columns are removed. The related model component is shown as table header. However, you can also return a single table:

```{r eval=successfully_loaded["glmmTMB"]}
mp <- model_parameters(model)
print(mp, split_component = FALSE)
```

## Adding model summaries

A model summary can be added to the table when `summary = TRUE` in the call to `model_parameters()`:

```{r}
model <- lm(Sepal.Length ~ Species * Petal.Length, data = iris)
model_parameters(model, summary = TRUE)
```

## Including the reference level of categorical variables

Sometimes, it can be helpful to include the reference level of categorical predictors in the table. This can be done by setting `include_reference = TRUE` (either directly in `model_parameters()` or in the `print()` method). Since the reference level is not a parameter, it is shown in a separate row, with `0` for the coefficient and blank cells for the remaining columns.

```{r}
model <- lm(Sepal.Length ~ Petal.Length + Species, data = iris)
model_parameters(model, include_reference = TRUE)
```

## Changing number of digits

`digits` changes the digits for coefficients, standard errors and statistics. `ci_digits` and `p_digits` are especially for the confidence intervals and p-values.

```{r}
model <- lm(Sepal.Length ~ Species, data = iris)
model_parameters(model, digits = 4)
```

p-values can be displayed in exact, scientific notation if required.

```{r}
model_parameters(model, p_digits = "scientific")
```

## Fixing column widths

By default, the width of table columns is set to the minimum required width. This works well for models that produce just one table. However, for models with multiple components, where each component is shown as separate table, columns are possibly no longer aligned across tables. See the following example from a zero-inflated mixed model that has three components (fixed count, fixed zero-inflated, random effects):

```{r eval=successfully_loaded["glmmTMB"]}
data("Salamanders")
# we create very long parameter names for this predictor here
levels(Salamanders$spp) <- paste("long", levels(Salamanders$spp))

model <- glmmTMB(
  count ~ spp + mined + (1 | site),
  ziformula = ~mined,
  family = poisson(),
  data = Salamanders
)

# default printing
model_parameters(model)
```

The `column_width` argument can be used to either define the width of specific columns, or to fix column widths of the same columns across tables to have the same width. In the latter case, use `column_width = "fixed"` in the `print()` method.

```{r eval=successfully_loaded["glmmTMB"]}
mp <- model_parameters(model)
print(mp, column_width = "fixed")
```

If `column_width` is a named vector, names are matched against column names, and those columns gain the specified minimum width.

```{r eval=successfully_loaded["glmmTMB"]}
print(mp, column_width = c(SE = 8, `95% CI` = 12, p = 7))
```

## Group parameters

The `groups` argument can be used to group parameters in the table. `groups` must be a named list, where the names of the list elements equal the header of each group, while the values of the list elements equal the parameter names, or the position of the parameters in the table (data frame).

In the following example, we see the names of the parameters in the `Parameter` column, while the rownumbers indicate their position.

```{r}
data(mtcars)
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$gear <- as.factor(mtcars$gear)
model <- lm(mpg ~ hp + gear * vs + cyl + drat, data = mtcars)

# don't select "Intercept" parameter
mp <- model_parameters(model, drop = "^\\(Intercept")

# inspect data frame
as.data.frame(mp)
```

Now we create a group named `"Engine"`, which encompasses the parameters `"cyl6"`, `"cyl8"`, `"vs"` and `"hp"`. The `"Interactions"` group includes `"gear4:vs"` and `"gear5:vs"`. The group `"controls"` has the parameters from rows 2, 3 and 7.

Note that the parameters in the table summary are re-ordered according to the order specified in `groups`.

```{r}
# group parameters, either by parameter name or position
print(mp, groups = list(
  "Engine" = c("cyl6", "cyl8", "vs", "hp"),
  "Interactions" = c("gear4:vs", "gear5:vs"),
  "Controls" = c(2, 3, 7)
)) # gear 4 and 5, drat
```

If you prefer tables without vertical borders, use the `sep` argument to define the string that is used as border-separator. This argument is passed down to `insight::export_table()`.

```{r}
# group parameters, either by parameter name or position
print(mp,
  sep = "  ",
  groups = list(
    "Engine" = c("cyl6", "cyl8", "vs", "hp"),
    "Interactions" = c("gear4:vs", "gear5:vs"),
    "Controls" = c(2, 3, 7)
  )
)
```

# Summaries for multiple models

`compare_parameters()` (or its alias `compare_models()`) allows to create tables for multiple models, aligned side by side.

By default, estimates and confidence intervals are shown.

```{r}
data(iris)
lm1 <- lm(Sepal.Length ~ Species, data = iris)
lm2 <- lm(Sepal.Length ~ Species + Petal.Length, data = iris)
lm3 <- lm(Sepal.Length ~ Species * Petal.Length, data = iris)
compare_parameters(lm1, lm2, lm3)
```

## Changing style of column output

By default, estimates and confidence intervals are shown. Using `select` allows us to create different output, e.g. standard errors instead of confidence intervals, or including p-values.

```{r}
compare_parameters(lm1, lm2, lm3, select = "se_p")
```

The `select` argument also has a basic support for glue-like syntax, which lets you layout model elements in a very flexible way. Following tokens are replaced by the related summary coefficients or statistics:

- `{estimate}` (or `{coefficient}` or `{coef}`): coefficient
- `{se}` (or `{std.error}` or `{standard error}`): standard error
- `{ci_low}` and `{ci_high}`: lower/upper confidence interval limits
- `{p}` (or `{pval}` or `{p.value}`): p-values
- `{stars}`: significant stars for p-values

Note that you have to add parentheses manually, e.g. around confidence intervals.

```{r}
# estimates, p-stars and standard error in parentheses
compare_parameters(lm1, lm2, lm3, select = "{estimate}{stars} ({se})")

# estimates, CI, p and stars
compare_parameters(lm1, lm2, lm3, select = "{estimate} ({ci_low}, {ci_high}), p={p}{stars}")
```

`select` also works for `model_parameters()`, however, it's necessary to call this argument via the `print()` method:

```{r}
# estimates, p-stars and CI in parentheses
result <- model_parameters(lm3)
print(result, select = "{estimate}{stars} ({ci})")
```


## Defining column names

The column names for the models are by default the objects' names. You can define own names using the `column_names` argument.

```{r}
compare_parameters(
  lm1, lm2, lm3,
  column_names = c("First Model", "Second Model", "Third Model")
)
```

## Models with multiple components

For models with multiple components, like mixed models with fixed and random effects, or models with count- and zero-inflation parts, using arguments `effects = "all"` and/or `component = "all"` prints separate tables for each model component.

```{r eval=successfully_loaded["glmmTMB"]}
library(glmmTMB)
data("fish")

m0 <- glm(count ~ child + camper, data = fish, family = poisson())

m1 <- glmmTMB(
  count ~ child + camper + (1 | persons) + (1 | ID),
  data = fish,
  family = poisson()
)

m2 <- glmmTMB(
  count ~ child + camper + zg + (1 | ID),
  ziformula = ~ child + (1 | persons),
  data = fish,
  family = truncated_poisson()
)

compare_parameters(m0, m1, m2, effects = "all", component = "all")
```

For such tables, they usually become clearer when the columns per model are aligned to a fixed width. You can do this using `column_width = "fixed"`.

```{r eval=successfully_loaded["glmmTMB"]}
cp <- compare_parameters(m0, m1, m2, effects = "all", component = "all")
print(cp, column_width = "fixed")
```


## Group parameters of multiple model tables

Grouping parameters works for `compare_parameters()` in the same way as shown above for `model_parameters()`.

**Note:** By default, the interaction mark is `×`, not `*` (see also section on global options [in this vignette](https://easystats.github.io/parameters/reference/print.parameters_model.html)). Since parameter names in `compare_parameters()` are already formatted before printing, `as.data.frame(cp)$Parameter` will probably return special unicode characters that you need to take care of in the `groups` argument (unless you use numeric indices).

```{r}
lm1 <- lm(Sepal.Length ~ Species + Petal.Length, data = iris)
lm2 <- lm(Sepal.Width ~ Species * Petal.Length, data = iris)

# remove intercept
cp <- compare_parameters(lm1, lm2, drop = "^\\(Intercept")

# look at parameters names, to know their names for "groups" argument
as.data.frame(cp)$Parameter # note the unicode char as interaction mark

# create groups. Interactions only present in 2nd model
print(cp, groups = list(
  Species = c(
    "Species (versicolor)",
    "Species (virginica)"
  ),
  Interactions = c(
    "Species (versicolor) × Petal Length", # note the unicode char!
    "Species (virginica) × Petal Length"
  ),
  Controls = "Petal Length"
))
```


# Splitting wide tables into multiple table parts

For very wide tables that cannot be displayed properly, you can use the `table_width` argument in the `print()` method to split tables into multiple parts. `table_width` can be a numeric value, or `"auto"`, indicating the width of the complete table. If `table_width = "auto"` and the table is wider than the current available width (i.e. line length) of the console (or any other source for textual output, like markdown files), the table is split into multiple parts. Else, if `table_width` is numeric and table rows are wider than `table_width`, the table is split into multiple parts.

```{r}
data(iris)
lm1 <- lm(Sepal.Length ~ Species, data = iris)
lm2 <- lm(Sepal.Length ~ Species + Petal.Length, data = iris)
lm3 <- lm(Sepal.Length ~ Species * Petal.Length, data = iris)
lm4 <- lm(Sepal.Length ~ Species * Petal.Length + Petal.Width, data = iris)

# very wide table
compare_parameters(lm1, lm2, lm3, lm4)

# table split into two parts
tab <- compare_parameters(lm1, lm2, lm3, lm4)
print(tab, table_width = 80)
```

# More advances tables and markdown / HTML formatting

The `print_md()` as well as `print_html()` functions can be used to create markdown (for knitting to PDF or Word) and HTML tables. 

Meanwhile, there are a lot of additional packages that allow users to have even more flexibility regarding table layouts. One package we can recommend is the [*modelsummary* package](https://vincentarelbundock.github.io/modelsummary/).
