---
title: "Robust Estimation of Standard Errors, Confidence Intervals, and p-values"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_width: 10.08
    fig_height: 6
tags: [r, parameters, variable selection, feature selection]
vignette: >
  %\VignetteIndexEntry{Robust Estimation of Standard Errors, Confidence Intervals, and p-values}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
options(knitr.kable.NA = "")
options(digits = 2)

knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  out.width = "100%"
)

if (!requireNamespace("dplyr", quietly = TRUE) ||
  !requireNamespace("clubSandwich", quietly = TRUE) ||
  !requireNamespace("sandwich", quietly = TRUE) ||
  !requireNamespace("lme4", quietly = TRUE)) {
  knitr::opts_chunk$set(eval = FALSE)
} else {
  library(parameters)
  library(dplyr)
}

set.seed(333)
```

The
[`model_parameters()`](https://easystats.github.io/parameters/articles/model_parameters.html)
function also allows the computation of standard errors, confidence intervals,
and *p*-values based on robust covariance matrix estimation from model
parameters. Robust estimation relies on
the [`sandwich`](https://cran.r-project.org/web/packages/sandwich/vignettes/sandwich-CL.pdf)
and
[`clubSandwich`](https://cran.r-project.org/web/packages/clubSandwich/index.html)
packages. This means that all models supported by either of these packages should work
with `model_parameters()` when `robust = TRUE`.

## Linear Regression Models

### Robust Covariance Matrix Estimation from Model Parameters

By default, when `model_parameters(robust = TRUE)`, it internally calls
`sandwich::vcovHC(type = "HC3")`. However, there are three arguments (see
[`?standard_error_robust`](https://easystats.github.io/parameters/reference/standard_error_robust.html)
for further details) that allow for choosing different methods and options of
robust estimation:
- `vcov_estimation`
- `vcov_type`
- `vcov_args`

Let us start with a simple example, which uses a heteroskedasticity-consistent
covariance matrix estimation with estimation-type "HC3" (i.e.
`sandwich::vcovHC(type = "HC3")`). 

First let's create a simple linear regression model, which we know violates
homoscedasticity assumption, and thus robust estimation methods are to be
considered.

```{r}
data(cars)
model <- lm(dist ~ speed, data = cars)

library(performance)
check_heteroscedasticity(model)
```

We would extract model parameters both with and without robust estimation to
highlight difference it makes to standard errors, confidence intervals, and
*p*-values. Also, note that the coefficient estimate and the *t*-statistic
associated with it remain unchanged.

```{r}
# model parameters, where SE, CI, and p-values are *not* based on robust estimation
model_parameters(model)

# model parameters, where SE, CI, and p-values are based on robust estimation
mp <- model_parameters(model, robust = TRUE)
mp

# compare standard errors to result from sandwich-package
mp$SE
unname(sqrt(diag(sandwich::vcovHC(model))))
```

### Cluster-Robust Covariance Matrix Estimation (sandwich)

If a different type of covariance matrix estimation is required, use the
`vcov_estimation`-argument. This argument needs the suffix for the related
`vcov*()`-functions as value, i.e. `vcov_estimation = "CL"` would call
`sandwich::vcovCL()`, or `vcov_estimation = "HAC"` would call
`sandwich::vcovHAC()`.

The specific estimation type can also be changed with `vcov_type`. For example,
`sandwich::vcovCL()` accepts estimation types `HC0` to `HC3`. In the next
example, we use a clustered covariance matrix estimation with `HC1`-estimation
type.

```{r}
# let's create a more complicated model
data(iris)
model <- lm(Petal.Length ~ Sepal.Length * Species + Sepal.Width, data = iris)

# change estimation-type
mp <- model_parameters(
  model,
  robust = TRUE,
  vcov_estimation = "CL", # type of covariance matrix
  vcov_type = "HC1" # type of robust estimation
)

mp

# compare standard errors to result from sandwich-package
mp$SE
unname(sqrt(diag(sandwich::vcovCL(model))))
```

Usually, clustered covariance matrix estimation is used when there is a
cluster-structure in the data. The variable indicating the cluster-structure can
be defined in `sandwich::vcovCL()` with the `cluster`-argument. In
`model_parameters()`, additional arguments that should be passed down to
functions from the `sandwich` package can be specified in `vcov_args`:

```{r}
iris$cluster <- factor(rep(LETTERS[1:8], length.out = nrow(iris)))

# change estimation-type, defining additional arguments
mp <- model_parameters(
  model,
  robust = TRUE,
  vcov_estimation = "CL",
  vcov_type = "HC1",
  vcov_args = list(cluster = iris$cluster)
)

mp

# compare standard errors to result from sandwich-package
mp$SE
unname(sqrt(diag(sandwich::vcovCL(model, cluster = iris$cluster))))
```

### Cluster-Robust Covariance Matrix Estimation (clubSandwich)

Cluster-robust estimation of the variance-covariance matrix can also be achieved
using `clubSandwich::vcovCR()`. Thus, when `vcov_estimation = "CR"`, the related
function from the `clubSandwich` package is called. Note that this function
_requires_ the specification of the `cluster`-argument.

```{r}
# create fake-cluster-variable, to demonstrate cluster robust standard errors
iris$cluster <- factor(rep(LETTERS[1:8], length.out = nrow(iris)))

# cluster-robust estimation
mp <- model_parameters(
  model,
  robust = TRUE,
  vcov_estimation = "CR",
  vcov_type = "CR1",
  vcov_args = list(cluster = iris$cluster)
)
mp

# compare standard errors to result from clubSsandwich-package
mp$SE

unname(sqrt(diag(clubSandwich::vcovCR(model, type = "CR1", cluster = iris$cluster))))
```

### Robust Covariance Matrix Estimation on Standardized Model Parameters

Finally, robust estimation can be combined with
[standardization](https://easystats.github.io/parameters/articles/model_parameters_standardized.html).
However, robust covariance matrix estimation only works for `standardize = "refit"`.

```{r}
# model parameters, robust estimation on standardized model
model_parameters(model, standardize = "refit", robust = TRUE)
```

## Linear Mixed-Effects Regression Models

### Robust Covariance Matrix Estimation for Mixed Models

For linear mixed-effects models, that by definition have a clustered
(*hierarchical* or *multilevel*) structure in the data, it is also possible to
estimate a cluster-robust covariance matrix. This is possible due to the
`clubSandwich` package, thus we need to define the same arguments as in the
above example.

```{r}
library(lme4)
data(iris)
set.seed(1234)
iris$grp <- as.factor(sample(1:3, nrow(iris), replace = TRUE))

# fit example model
model <- lme4::lmer(
  Sepal.Length ~ Species * Sepal.Width + Petal.Length + (1 | grp),
  data = iris
)

# model parameters without robust estimation
model_parameters(model)

# model parameters with cluster robust estimation
model_parameters(
  model,
  robust = TRUE,
  vcov_estimation = "CR",
  vcov_type = "CR1",
  vcov_args = list(cluster = iris$grp)
)
```

Notice that robust estimation returns different standard errors, confidence
intervals, and *p*-values compared to the standard estimation. Also, note that
the coefficient estimate and the statistic associated with it remain unchanged.

### Robust Covariance Matrix Estimation on Standardized Mixed Model Parameters

Once again, robust estimation can be combined with standardization for linear
mixed-effects models as well and works only with `standardize = "refit"`.

```{r}
# model parameters, cluster robust estimation of standardized mixed model
model_parameters(
  model,
  standardize = "refit",
  robust = TRUE,
  vcov_estimation = "CR",
  vcov_type = "CR1",
  vcov_args = list(cluster = iris$grp)
)
```

Notice how drastically some of the *p*-values change between
robust-unstandardized model and robust-standardized model.

<!-- TO DO: Maybe provide references to read more? -->