---
title: "Bootstrapped parameters"
output: 
  github_document:
    toc: true
    fig_width: 10.08
    fig_height: 6
  rmarkdown::html_vignette:
    toc: true
    fig_width: 10.08
    fig_height: 6
tags: [r, bayesian, gam, smooth]
vignette: >
  %\VignetteIndexEntry{Bootstrapped parameters}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
options(knitr.kable.NA = '')
knitr::opts_chunk$set(comment=">")
options(digits=2)

if (!requireNamespace("ggplot2", quietly = TRUE) ||
  !requireNamespace("parameters", quietly = TRUE) ||
  !requireNamespace("poorman", quietly = TRUE) ||
  !requireNamespace("tidyr", quietly = TRUE)) {
  knitr::opts_chunk$set(eval = FALSE)
} else {
  library(parameters)
library(ggplot2)
library(poorman)
library(tidyr)
}

set.seed(333)
```


The basic idea of bootstrapping is that inference about a parent population from sample data can be modelled by resampling the sample data. It is often used as an alternative to statistical inference based on the assumption of a parametric model when that assumption is in doubt, or where parametric inference is complicated.

## Comparison between regular and boostrapped estimates

### Data

In order to compare point-estimates with bootstrapped parameters for frequentist models. We generated one large sample (the **parent population**, size `1000000`) of two continuous variables producing a regression coefficient of `0.5`. We then iteratively extracted a subsample of size `30`, computed 3 types of coefficient (regular, bootstrapped median with 1000 and 4000 iterations) that were substracted from the "parent" coefficient. The closer the value is from 0, and the closer it is from the "true" effect.

The data is available on githuband the code to generate it is available [here](https://easystats.github.io/circus/articles/bootstrapped.html).


```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(poorman)
library(tidyr)

df <- read.csv("https://raw.github.com/easystats/circus/master/data/bootstrapped.csv")
```


### Visualisation
```{r message=FALSE, warning=FALSE, fig.height=10, fig.width=8}
library(see)

df_long <- df %>%
  select(Coefficient, Bootstrapped_1000, Bootstrapped_4000) %>%
  gather(Type, Distance) %>%
  mutate(Type = forcats::fct_relevel(Type, c("Coefficient", "Bootstrapped_1000", "Bootstrapped_4000")))

df_long %>% 
  ggplot(aes(y = Distance, x = Type, fill = Type)) +
  geom_violin() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = c("#2196F3", "#FF9800", "#f44336")) +
  theme_modern() +
  ylab("Distance (0 is the parent / true effect)")
```


### Testing

#### Bayes factor analysis
```{r message=FALSE, warning=FALSE}
library(BayesFactor)
library(bayestestR)

bayestestR::bayesfactor(BayesFactor::ttestBF(df$Coefficient))
bayestestR::bayesfactor(BayesFactor::ttestBF(df$Bootstrapped_1000))
bayestestR::bayesfactor(BayesFactor::ttestBF(df$Bootstrapped_4000))
```

#### Contrast analysis
```{r message=FALSE, warning=FALSE}
library(emmeans)

lm(Distance ~ Type, data = df_long) %>% 
  emmeans::emmeans(~Type)
```


### Conclusion 

Negligible difference, but bootstrapped (n=1000) seems (very) slightly more accurate.
