---
title: "Robust Estimation of Standard Errors, Confidence Intervals and p-values"
output:
  github_document:
    toc: true
    fig_width: 10.08
    fig_height: 6
  rmarkdown::html_vignette:
    toc: true
    fig_width: 10.08
    fig_height: 6
tags: [r, parameters, variable selection, feature selection]
vignette: >
  %\VignetteIndexEntry{Robust Estimation of Standard Errors}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
bibliography: bibliography.bib
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
options(knitr.kable.NA = "")
options(digits = 2)
knitr::opts_chunk$set(comment = "#>")

if (!requireNamespace("poorman", quietly = TRUE) ||
  !requireNamespace("clubSandwich", quietly = TRUE) ||
  !requireNamespace("sandwich", quietly = TRUE) ||
  !requireNamespace("lme4", quietly = TRUE)) {
  knitr::opts_chunk$set(eval = FALSE)
} else {
  library(parameters)
  library(poorman)
  library(clubSandwich)
  library(lme4)
}

set.seed(333)
```

The [`model_parameters()`](https://easystats.github.io/parameters/articles/model_parameters.html) function also allows the computation of standard errors, confidence intervals and p-values based on robust covariance matrix estimation from model parameters. Robust estimation is based on the packages **sandwich** and **clubSandwich**, so all models supported by either of these packages work with `model_parameters()`.

There are two arguments responsible to define the heteroscedasticity-consistent covariance matrix, that should be used for robust estimation: `vcov` and `vcov_args`.

The `vcov` argument defines the variance-covariance matrix used to compute uncertainty estimates. This argument accepts following inputs:

* A covariance matrix
* A function which returns a covariance matrix (e.g., `stats::vcov()`)
* A string which indicates the kind of uncertainty estimates to return.
  - Heteroskedasticity-consistent: `"HC"`, `"HC0"`, `"HC1"`, `"HC2"`, `"HC3"`, `"HC4"`, `"HC4m"`, `"HC5"`. See `?sandwich::vcovHC`
  - Cluster-robust: `"CR"`, `"CR0"`, `"CR1"`, `"CR1p"`, `"CR1S"`, `"CR2"`, `"CR3"`. See `?clubSandwich::vcovCR`
  - Bootstrap: `"BS"`, `"xy"`, `"residual"`, `"wild"`, `"mammen"`, `"fractional"`, `"jackknife"`, `"norm"`, `"webb"`. See `?sandwich::vcovBS`
  - Other `sandwich` package functions: `"HAC"`, `"PC"`, `"vCL"`, `"PL"`.


## Classical Regression Models

### Robust Covariance Matrix Estimation from Model Parameters

The simplest option, `model_parameters(vcov = "HC)`, internally calls `sandwich::vcovHC(type = "HC3")`. However, `vcov` and `vcov_args` can be used for different options to calculate the heteroscedasticity-consistent covariance matrix (see [`?standard_error_robust`](https://easystats.github.io/parameters/reference/standard_error_robust.html) for further details).

Let us start with a simple example, which uses a heteroskedasticity-consistent covariance matrix estimation with estimation-type "HC3" (i.e. `sandwich::vcovHC(type = "HC3")` is called):

```{r}
data(iris)
model <- lm(Petal.Length ~ Sepal.Length * Species + Sepal.Width, data = iris)

# model parameters, where SE, CI and p-values are based on robust estimation
mp <- model_parameters(model, vcov = "HC")
mp

# compare standard errors to result from sandwich-package
mp$SE
unname(sqrt(diag(sandwich::vcovHC(model))))
```

### Cluster-Robust Covariance Matrix Estimation (sandwich)

The specific estimation type can be either passed as string to `vcov`, or changed with the `type` option that is passed to the `vcov_args` argument:

* `vcov = "HC1"`
* `vcov = "HC", vcov_args = list(type = "HC1")`.

If the `type` is acceptes by different functions, e.g., both `sandwich::vcovHC()` and `sandwich::vcovCL()` accepts estimation types HC0 to HC3, we need the "long" form to specify all relevant options. In the next example, we use a clustered covariance matrix estimation with HC1-estimation type.

```{r}
# change estimation-type
mp <- model_parameters(model, vcov = "CL", vcov_args = list(type = "HC2"))
mp

# compare standard errors to result from sandwich-package
mp$SE
unname(sqrt(diag(sandwich::vcovCL(model, type = "HC2"))))
```

Usually, clustered covariance matrix estimation is used when there is a cluster-structure in the data. The variable indicating the cluster-structure can be defined in `sandwich::vcovCL()` with the `cluster`-argument. As mentioned above, in `model_parameters()`, additional arguments that should be passed down to functions from the **sandwich** package can be specified in `vcov_args`:

```{r}
iris$cluster <- factor(rep(LETTERS[1:8], length.out = nrow(iris)))
# change estimation-type, defining additional arguments
mp <- model_parameters(
  model,
  vcov = "CL",
  vcov_args = list(type = "HC2", cluster = iris$cluster)
)
mp

# compare standard errors to result from sandwich-package
mp$SE
unname(sqrt(diag(sandwich::vcovCL(model, type = "HC2", cluster = iris$cluster))))
```

### Cluster-Robust Covariance Matrix Estimation (clubSandwich)

Cluster-robust estimation of the variance-covariance matrix can also be achieved using `clubSandwich::vcovCR()`. Thus, when `vcov = "CR"`, the related function from the **clubSandwich** package is called. Note that this function _requires_ the specification of the `cluster`-argument.

```{r}
# create fake-cluster-variable, to demonstrate cluster robust standard errors
iris$cluster <- factor(rep(LETTERS[1:8], length.out = nrow(iris)))

# cluster-robust estimation
mp <- model_parameters(
  model,
  vcov = "CR",
  vcov_args = list(type = "CR1", cluster = iris$cluster)
)
mp

# compare standard errors to result from clubSsandwich-package
mp$SE
unname(sqrt(diag(clubSandwich::vcovCR(model, type = "CR1", cluster = iris$cluster))))
```


### Robust Covariance Matrix Estimation on Standardized Model Parameters

Finally, robust estimation can be combined with standardization. However, robust covariance matrix estimation only works for `standardize = "refit"`.

```{r}
# model parameters, robust estimation on standardized model
model_parameters(model, standardize = "refit", vcov = "HC")
```


## Mixed Models

### Robust Covariance Matrix Estimation for Mixed Models

For linear mixed models, that by definition have a clustered ("hierarchical" or multilevel) structure in the data, it is also possible to estimate a cluster-robust covariance matrix. This is possible due to the **clubSandwich** package, thus we need to define the same arguments as in the above example.

```{r}
library(lme4)
data(iris)
set.seed(1234)
iris$grp <- as.factor(sample(1:3, nrow(iris), replace = TRUE))

# fit example model
model <- lme4::lmer(
  Sepal.Length ~ Species * Sepal.Width + Petal.Length + (1 | grp),
  data = iris
)

# normal model parameters, like from 'summary()'
model_parameters(model)

# model parameters, cluster robust estimation for mixed models
model_parameters(
  model,
  vcov = "CR",
  vcov_args = list(type = "CR1", cluster = iris$grp)
)
```

### Robust Covariance Matrix Estimation on Standardized Mixed Model Parameters

Again, robust estimation can be combined with standardization for linear mixed models as well, which in such cases also only works for `standardize = "refit"`.

```{r}
# model parameters, cluster robust estimation on standardized mixed model
model_parameters(
  model,
  standardize = "refit",
  vcov = "CR",
  vcov_args = list(type = "CR1", cluster = iris$grp)
)
```
